Dynamical modeling of tidal streams

### 2014-01-06 tex/ms.tex
streams: Scott's comments

The distribution of frequencies is three dimensional. Because we model
the leading and the trailing arm separately, a single-peaked
distribution suffices to describe the frequency distribution. A
general description in terms of a Gaussian distribution would have
nine free parameters. However, we can model it with fewer free
parameters as follows. First, following \citet{Eyre11a}, we
approximate the \emph{action} distribution as a Gaussian with standard
deviations
\begin{align}
  \sigma_{J_R} & = \frac{1}{2\,\pi}\oint\dd r\,p_r \approx \frac{1}{\pi}\,\sigv\,\left(r_{\mathrm{apo}}-r_{\mathrm{peri}}\right)\,,\\
  \sigma_{L_Z} & \approx  \sigv\,r_{\mathrm{peri}}\,,\\
  \sigma_{J_Z} & = \frac{1}{\pi}\oint\dd z\,p_z \approx \frac{2}{\pi}\,\sigv\,Z_{\mathrm{max}}\,,
\end{align}
where $\sigv$ is the velocity dispersion of the progenitor,
$r_{\mathrm{apo}}$ and $r_{\mathrm{peri}}$ are the apo- and pericenter
radii of the progenitor orbit, respectively, and $Z_{\mathrm{max}}$ is
the maximum $Z$ reached on this orbit. We assume that all of the
correlations between the actions are zero (see
\figurename~\ref{fig:gd1_jt}). Then we propagate this Gaussian
variance to frequency space using the Hessian $(\partial \veco /
\partial \vecj)\large|_{\vecj^p}$\footnote{The Hessian $(\partial
  \veco / \partial \vecj)\large|_{\vecj^p}$ can be computed using the
  method in \appendixname~\ref{sec:aa} by computing $\partial
  (\vecj,\veca) / \partial (\vecx,\vecv)$ and $\partial (\veco,\veca)
  / \partial (\vecx,\vecv)$ and forming $\partial \veco / \partial
  \vecj = \left(\partial (\vecj,\veca) / \partial
  (\vecx,\vecv)\right)^{-1}\,\left(\partial (\veco,\veca) / \partial
  (\vecx,\vecv)\right)$.}, which gives the variance matrix
$V_\veco$. The principal eigenvector of this matrix is the model
direction along which the stream spreads. While the model I propose
here is a simple one, it works well without having been tweaked for
the simulation under scrutiny. The angle between the principal
eigenvector of $V_\veco$ and the progenitor's frequency vector is
$0.50^\circ$ while the value measured from the simulation is
$0.18^\circ$. For a model with an isotropic action distribution, the
misalignment would be $1.28^\circ$.

@@

The forward model of this paper is superior to all other
stream-fitting methods in that it allows easy marginalization over
missing or noisily measured phase--space coordinates. The obvious
exception to this is orbit fitting, which as a purely one-dimensional
model is easy to integrate over. However, orbit fitting likely
produces biased results as streams do not delineate single orbits over
their full angle widths (see below). The fiducial model I propose in
this paper allows for a straightforward replacement of orbit fitting,
at the expense of adding two extra parameters, $\sigv$ and $t_d$, in
addition to the six-dimensional position of the progenitor. The orbit
in orbit fitting then gets replaced by the stream track, calculated as
in \sectionname~\ref{sec:track}. To constrain $\sigv$ and $t_d$ it is
necessary to also use the observed width and length of the
stream. While cold streams are typically too narrow to be resolved in
$D$ or any of the velocity components, the observed width of the
stream's sky projection can be measured and used as a
constraint. Future high-resolution spectroscopic data,
$\mu\mathrm{as}$-level astrometric data, or percent-level distances
for standard candles may allow the width and structure of tidal
streams to be resolved and provide further constraints on the
generative model. The switch from orbit fitting to stream-track
fitting comes at the expense of $\approx100$ orbit integrations for
each model rather than a single integration. If the progenitor is
unknown we also need to marginalize over whether we are seeing a
leading or a trailing stream.

@@

While this requires high-quality data, this should be achievable in
the near future. We can average or ``stack'' noisily measured
phase--space coordinates in small segments of the stream to determine
high quality average phase--space positions and use these instead of
data on individual stream stars. These considerations show that
further observations of members of the GD-1 and Orphan streams would
be highly informative for determining the location or fate of their
progenitors.

>>>>

The distribution of frequencies is three dimensional. Because we model
the leading and the trailing arm separately, a single-peaked
distribution suffices to describe the frequency distribution. A
general description in terms of a Gaussian distribution would have
nine free parameters. However, we can model it with fewer free
parameters as follows. First, following \citet{Eyre11a}, we
approximate the \emph{action} distribution as a Gaussian with standard
deviations given by the approximate spread in the actions $(\delta
J_R,\delta L_Z,\delta J_Z)$ in the cluster
\begin{align}
  \sigma_{J_R} & \approx \delta \left(\frac{1}{2\,\pi}\oint\dd r\,p_r\right) \approx \frac{1}{\pi}\,\sigv\,\left(r_{\mathrm{apo}}-r_{\mathrm{peri}}\right)\,,\\
  \sigma_{L_Z} & \approx  \delta L_Z \approx \sigv\,r_{\mathrm{peri}}\,,\\
  \sigma_{J_Z} & \approx \delta\left(\frac{1}{\pi}\oint\dd z\,p_z \right)\approx \frac{2}{\pi}\,\sigv\,Z_{\mathrm{max}}\,,
\end{align}
where $\sigv$ is the velocity dispersion of the progenitor,
$r_{\mathrm{apo}}$ and $r_{\mathrm{peri}}$ are the apo- and pericenter
radii of the progenitor orbit, respectively, and $Z_{\mathrm{max}}$ is
the maximum $Z$ reached on this orbit. We assume that all of the
correlations between the actions are zero (see
\figurename~\ref{fig:gd1_jt}). Then we propagate this Gaussian
variance to frequency space using the Hessian $(\partial \veco /
\partial \vecj)\large|_{\vecj^p}$\footnote{The Hessian $(\partial
  \veco / \partial \vecj)\large|_{\vecj^p}$ can be computed using the
  method in \appendixname~\ref{sec:aa} by computing $\partial
  (\vecj,\veca) / \partial (\vecx,\vecv)$ and $\partial (\veco,\veca)
  / \partial (\vecx,\vecv)$ and forming $\partial \veco / \partial
  \vecj = \left(\partial (\vecj,\veca) / \partial
  (\vecx,\vecv)\right)^{-1}\,\left(\partial (\veco,\veca) / \partial
  (\vecx,\vecv)\right)$.}, which gives the variance matrix
$V_\veco$. The principal eigenvector of this matrix is the model
direction along which the stream spreads. While the model I propose
here is a simple one, it works well without having been tweaked for
the simulation under scrutiny. The angle between the principal
eigenvector of $V_\veco$ and the progenitor's frequency vector is
$0.50^\circ$ which is close to the value measured from the simulation
is $0.18^\circ$. For a model with an isotropic action distribution,
the misalignment would be $1.28^\circ$, much larger than the measured
value.

@@

The forward model of this paper is superior to other stream-fitting
methods in that it allows easy marginalization over missing or noisily
measured phase--space coordinates. The obvious exception to this is
orbit fitting, which as a purely one-dimensional model is easy to
integrate over. However, orbit fitting likely produces biased results
as streams do not delineate single orbits over their full angle widths
(see below). The fiducial model I propose in this paper allows for a
straightforward replacement of orbit fitting, at the expense of adding
two extra parameters, $\sigv$ and $t_d$, in addition to the
six-dimensional position of the progenitor. The orbit in orbit fitting
then gets replaced by the stream track, calculated as in
\sectionname~\ref{sec:track}. To constrain $\sigv$ and $t_d$ it is
necessary to also use the observed width and length of the
stream. While cold streams are typically too narrow to be resolved in
$D$ or any of the velocity components, the observed width of the
stream's sky projection can be measured and used as a
constraint. Future high-resolution spectroscopic data,
$\mu\mathrm{as}$-level astrometric data, or percent-level distances
for standard candles may allow the width and structure of tidal
streams to be resolved and provide further constraints on the
generative model. The switch from orbit fitting to stream-track
fitting comes at the expense of $\approx100$ orbit integrations for
each model rather than a single integration. If the progenitor is
unknown we also need to marginalize over whether we are seeing a
leading or a trailing stream.

@@

While determining the mean orbit as a function of position along the
stream requires high-quality data, this should be achievable in the
near future. We can average or ``stack'' noisily measured phase--space
coordinates in small segments of the stream to determine high quality
average phase--space positions and use these instead of data on
individual stream stars. These considerations show that further
observations of members of the GD-1 and Orphan streams would be highly
informative for determining the location or fate of their progenitors.
